Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2015-09-10 draft.tex
change 'n' to 't' throughout appendix

In Section~\ref{sec:mainproof}, we considered the outside actors' abilities to make investments to increase the various payoffs of the home state government and the secessionists. The international community, in particular, often employs another option by joining the home state in enforcing economic sanctions against the unrecognized state, an action that \emph{reduces} the secessionists' payoffs from the status quo (i.e. sanctions are equivalent to $R_{cn} < 0$). Note that this may be particularly effective if $c$ is a large coalition of states acting in concert.

@@

{\bf Proof of Lemma 1} \\
Recall that the upper case notation represents payoffs gross of investments by the outside players. Without loss of generality, consider the incentives to deviate for player $s$ in period $n$ given that the other three players play their equilibrium actions. Given the stationary equilibrium actions, playing Status Quo will lead to $Q_{sn} \ \forall n$, so the continuation value is $\frac{Q_{sn}}{1-\delta}$. Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_{sn}$ in each period, is $\frac{L_{sn}}{1-\delta}$. The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, and $L_s$ forever with probability $1-p_1$. Thus the continuation value is $\zeta_s +\frac{W_{sn}(p_1)+ L_{sn}(1-p_1)}{1-\delta}$.

In order for player $s$ to play Status Quo, it must be that the continuation value from Status Quo is higher than both that from playing Cede (Condition 1) and Fight (Condition 2). The argument for player $g$ is symmetric. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 2} \\
In the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$. If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_{pn}^{RU}= -\alpha -R_p^{RU}$ and $U_{cn}^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.

@@

Player $c$ would be able to deter player $p$ from investing to maintain the status quo---e.g. counter an addition to $q_{sn}$ with an investment to $l_{sn}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} \geq Q_{sn} - l_{sn}$, where $Q_{sn}$ is gross of any investment by $p$.\footnote{Here we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting. If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the gain player $c$ receives from investing.

Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{\alpha}{1 -\de} > \frac{\beta+ R_c^{SQ}}{1 -\de}$, $p$ is willing to invest enough to create a large enough difference in $Q_{sn} - l_{sn}$ so that player $c$ will not find contesting the status quo to be in its interest. Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$. Thus a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome is $\alpha  > \beta$. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 3} \\
Again, in the status quo, we have $U_{pn}^{SQ}= -R_p^{SQ}$ and $U_{cn}^{SQ}= -R_c^{SQ}$. If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pn}^{REC}= \lambda -R_p^{REC}$ and $U_{cn}^{REC}= -\nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.

Thus the difference in continuation values from switching from the status quo to recognition for player $p$ is $\frac{\lambda +R_p^{SQ}}{1 -\de} -R_p^{REC}$. For player $c$ it is $\frac{-\nu + R_c^{SQ}}{1 -\de}-R_c^{REC}$. $R_c^{REC} =0$ since the international community will not invest to create its least preferred outcome. So $c$ is willing to invest up to $\nu$ per period toward maintaining the status quo, and $p$ is willing to invest up to $\frac{\lambda +R_p^{SQ}}{1 -\de}$ toward recognition. We have shown that the largest possible investments by $p$ in the status quo are $\frac{\beta}{1 -\de}$ in $n=1$ and $\mu$ in each period thereafter, so the largest this difference can be for $p$ is $\frac{\lambda + \mu + \beta}{1 -\de}$.

Player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g. counter an addition to $l_{gn}$ with an investment to $q_{gn}$---as long as $\nu \geq L_{gn} - q_{gn}$, where $L_{gn}$ is gross of any investment by $p$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $n$ is such that ceding is better than fighting. If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the loss player $c$ receives from investing.

Given our assumption that the status quo outcome is no worse than the recognition outcome net of investments for player $g$, $p$'s willingness to invest up to $\frac{\lambda + \mu + \beta}{1 -\de}$ can create a difference of at most that amount in $L_{gn} - q_{gn}$.

@@

Multiplying Condition 3 by $-(1-p_{1s})$, we have $-\beta \cdot (1-p_{1s}) \geq -\alpha \cdot (1-p_{1s})$. Using this with Condition 4, we have $\nu p_{1s} - \beta (1-p_{1s}) \geq \lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta> 0 \ \forall p_{1s}$ and therefore also
\[
  \textstyle -\left(\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}\right) \geq \frac{\lambda p_{1s} - \alpha (1-p_{1s})+ \mu + \beta}{1 - \de}> 0 \ \forall p_{1s}
\]

Hence when player $p$'s gain from instigating war is positive, player $c$'s loss is negative and larger in absolute magnitude. Player $c$ will thus counter any such investment by player $p$, deterring the investment in the first place. \hfill $\blacksquare$
\\
\\
{\bf Proof of Proposition 2} \\
Suppose that sanctions are applied by player $c$ in period $n+1$ after a Status Quo equilibrium such as the one supported by the construction in Proposition 1 has been established in period $n$. As the only alteration is a reduction in $Q_{sn}$ by the amount of the sanctions, it is immediate that if player $p$ is willing and able to supplement player $s$'s status quo payoffs to counteract sanctions, then player $s$'s incentives to play Status Quo are not disturbed. As in Proposition 1, it is not necessary that the patron make these payments immediately if player $s$'s utility from playing Status Quo is high enough relative to its other options, but the degradation of the status quo payoffs will eventually make these payments necessary in order to sustain a Status Quo equilibrium.

>>>>

In Section~\ref{sec:mainproof}, we considered the outside actors' abilities to make investments to increase the various payoffs of the home state government and the secessionists. The international community, in particular, often employs another option by joining the home state in enforcing economic sanctions against the unrecognized state, an action that \emph{reduces} the secessionists' payoffs from the status quo (i.e. sanctions are equivalent to $R_{ct} < 0$). Note that this may be particularly effective if $c$ is a large coalition of states acting in concert.

@@

{\bf Proof of Lemma 1} \\
Recall that the upper case notation represents payoffs gross of investments by the outside players. Without loss of generality, consider the incentives to deviate for player $s$ in period $t$ given that the other three players play their equilibrium actions. Given the stationary equilibrium actions, playing Status Quo will lead to $Q_{st} \ \forall t$, so the continuation value is $\frac{Q_{st}}{1-\delta}$. Player $s$'s continuation value from the one-shot deviation to Cede, which leads to an absorbing state with payoff $L_{st}$ in each period, is $\frac{L_{st}}{1-\delta}$. The one-shot deviation to fight results in the war lottery, with a cost of $-\zeta_s$, $W_s$ forever with probability $p_1$, and $L_s$ forever with probability $1-p_1$. Thus the continuation value is $\zeta_s +\frac{W_{st}(p_1)+ L_{st}(1-p_1)}{1-\delta}$.

In order for player $s$ to play Status Quo, it must be that the continuation value from Status Quo is higher than both that from playing Cede (Condition 1) and Fight (Condition 2). The argument for player $g$ is symmetric. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 2} \\
In the status quo, we have $U_{pt}^{SQ}= -R_p^{SQ}$ and $U_{ct}^{SQ}= -R_c^{SQ}$. If the outcome is reunification, payoffs in this absorbing state in the period in which the investment is made are $U_{pt}^{RU}= -\alpha -R_p^{RU}$ and $U_{ct}^{RU}= \beta -R_c^{RU}$ where $SQ$ and $RU$ distinguish investments under the status quo and reunification scenarios respectively.

@@

Player $c$ would be able to deter player $p$ from investing to maintain the status quo---e.g. counter an addition to $q_{st}$ with an investment to $l_{st}$---as long as $R_c^{RU} = \frac{\beta + R_c^{SQ}}{1 -\de} \geq Q_{st} - l_{st}$, where $Q_{st}$ is gross of any investment by $p$.\footnote{Here we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting. If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the gain player $c$ receives from investing.

Given our assumption that the status quo outcome is no worse than the reunification outcome net of investments for player $s$, when $\frac{\alpha}{1 -\de} > \frac{\beta+ R_c^{SQ}}{1 -\de}$, $p$ is willing to invest enough to create a large enough difference in $Q_{st} - l_{st}$ so that player $c$ will not find contesting the status quo to be in its interest. Whenever the status quo will be the outcome in this way, $c$ maximizes its utility by choice of $R_c^{SQ} =0$. Thus a sufficient condition to prevent $c$ from inducing the \emph{reunification} outcome is $\alpha  > \beta$. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 3} \\
Again, in the status quo, we have $U_{pt}^{SQ}= -R_p^{SQ}$ and $U_{ct}^{SQ}= -R_c^{SQ}$. If the outcome is recognition of the secessionists, per-period payoffs in this absorbing state become $U_{pt}^{REC}= \lambda -R_p^{REC}$ and $U_{ct}^{REC}= -\nu -R_c^{REC}$ where $SQ$ and $REC$ distinguish investments under the status quo and recognition scenarios respectively.

Thus the difference in continuation values from switching from the status quo to recognition for player $p$ is $\frac{\lambda +R_p^{SQ}}{1 -\de} -R_p^{REC}$. For player $c$ it is $\frac{-\nu + R_c^{SQ}}{1 -\de}-R_c^{REC}$. $R_c^{REC} =0$ since the international community will not invest to create its least preferred outcome. So $c$ is willing to invest up to $\nu$ per period toward maintaining the status quo, and $p$ is willing to invest up to $\frac{\lambda +R_p^{SQ}}{1 -\de}$ toward recognition. We have shown that the largest possible investments by $p$ in the status quo are $\frac{\beta}{1 -\de}$ in $t=1$ and $\mu$ in each period thereafter, so the largest this difference can be for $p$ is $\frac{\lambda + \mu + \beta}{1 -\de}$.

Player $c$ would be willing to counter whatever investment player $p$ makes to try to induce recognition---e.g. counter an addition to $l_{gt}$ with an investment to $q_{gt}$---as long as $\nu \geq L_{gt} - q_{gt}$, where $L_{gt}$ is gross of any investment by $p$.\footnote{As in the proof of Lemma 2, we have assumed that the payoff relationship at $t$ is such that ceding is better than fighting. If this relationship is reversed, similar analysis and results hold.} That is, the required investment must be no more than the loss player $c$ receives from investing.

Given our assumption that the status quo outcome is no worse than the recognition outcome net of investments for player $g$, $p$'s willingness to invest up to $\frac{\lambda + \mu + \beta}{1 -\de}$ can create a difference of at most that amount in $L_{gt} - q_{gt}$.

@@

Multiplying Condition 3 by $-(1-p_{1s})$, we have $-\beta \cdot (1-p_{1s}) \geq -\alpha \cdot (1-p_{1s})$. Using this with Condition 4, we have $\nu p_{1s} - \beta (1-p_{1s}) \geq \lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta> 0 \ \forall p_{1s}$ and therefore also
\[
  \textstyle -\left(\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}\right) \geq \frac{\lambda p_{1s} - \alpha (1-p_{1s})+ \mu + \beta}{1 - \de}> 0 \ \forall p_{1s}.
\]

Hence when player $p$'s gain from instigating war is positive, player $c$'s loss is negative and larger in absolute magnitude. Player $c$ will thus counter any such investment by player $p$, deterring the investment in the first place. \hfill $\blacksquare$
\\
\\
{\bf Proof of Proposition 2} \\
Suppose that sanctions are applied by player $c$ in period $t+1$ after a Status Quo equilibrium such as the one supported by the construction in Proposition 1 has been established in period $t$. As the only alteration is a reduction in $Q_{st}$ by the amount of the sanctions, it is immediate that if player $p$ is willing and able to supplement player $s$'s status quo payoffs to counteract sanctions, then player $s$'s incentives to play Status Quo are not disturbed. As in Proposition 1, it is not necessary that the patron make these payments immediately if player $s$'s utility from playing Status Quo is high enough relative to its other options, but the degradation of the status quo payoffs will eventually make these payments necessary in order to sustain a Status Quo equilibrium.
